---
title: "Almond Yield as a Function"
subtitle: "EDS 230 - Assignment #2"
date: "04/14/2025"
format: html
author: "Jordan Sibley and Marina Kochuten" 
editor_options: 
  chunk_output_type: console
---

## Overview 

The purpose of this document is to practice building models and utilizing functions for environmental data. The goal is to implement a simple model of almond yield anomaly response to climate variables. The source of the function comes from the study, *Impacts of future climate change on California perennial crop yields: Model projections with climate and crop uncertainties* (Lobell et al. 2006). 

## Model Set Up 

### Inputs & Outputs 

- **Inputs**: daily times series of minimum, maximum daily temperatures and precipitation
- **Outputs**: maximum, minimum and mean yield anomaly for an input time series

### Model diagram 

INSERT DIAGRAM 


## Function Overview 



### Load packages and data 
```{r}
# Load libraries
library(tidyverse)
library(here)

# Read in climate data
data <- read.table("data/clim.txt", header = TRUE, sep = " ", stringsAsFactors = FALSE)
```


## Running the Model 

#### Preparing dataset for function input 
```{r}
# Filter not complete years - no Jan or Feb data for 1988
data_clean <- data |> 
  filter(year != 1988)
```

#### Apply function to test climate data
```{r}
# Calculate almost yield anomalies
calculate_almond_yield(data_clean)
```



