---
title: "Almond Yield as a Function"
subtitle: "EDS 230 - Assignment #2"
date: "04/14/2025"
format: html
author: "Jordan Sibley and Marina Kochuten" 
---

## Overview 

The purpose of this document is to practice building models and utilizing functions for environmental data. The goal is to implement a simple model of almond yield anomaly response to climate variables. The source of the function comes from the study, *Impacts of future climate change on California
perennial crop yields: Model projections with
climate and crop uncertainties* (Lobell et al. 2006). 

## Model Set Up 

### Inputs & Outputs 

- **Inputs**: daily times series of minimum, maximum daily temperatures and precipitation
- **Outputs**: maximum, minimum and mean yield anomoly for a input time series

### Model diagram 

INSERT DIAGRAM 


## Function Overview 

```{r}
# test workflow to translate to function 

# Aggregate daily climate data to monthly averages
df_monthly <- data |>
  group_by(year, month) |>
  summarise(tmax_c_avg = mean(tmax_c),
            tmin_c_avg = mean(tmin_c),
            precip_avg = mean(precip))

# 1988 and 2010 are not a full years - exclude from data
df_monthly <- df_monthly |>
  filter(year != 1988 & year != 2010)
  
# calculate annual yield
# I want a way for the code to look through the dataframe and pull out the correct tmin, tmax, and precip values for each year
# The equation uses min and max temp for month 2, and precip for month 1



```


## Running the Model 

### Load packages and data 
```{r}
library(tidyverse)
library(here)

# Climate data 
data <- read.table("data/clim.txt", header = TRUE, sep = " ", stringsAsFactors = FALSE)
```

